---
title: "aqs"
author: "Ian Jiang"
date: "November 24, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Loading in necessary packages}
install.packages("MASS")
library(MASS)

install.packages("faraway")
library(faraway)

install.packages("nlme")
library(nlme)
```



```{r Loading data}
#Loading data
data <- read.csv("data_nomiss.csv")

#Dropping missing values in response
data <- data[!is.na(data$AQI),]

#Assigning date from year, month, day
data$date <- as.numeric( difftime(ISOdate(data$Year, data$Month, data$Day, 0), ISOdate(2016,1,1,0), units="days"))
```

```{r Fitting full model} 

#Fitting full model
full.lmod <- lm(formula = AQI ~ Nitrous.Oxide + County + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data)

```

```{r Diagnostics}
#--------------------Constant Variance--------------------

#Plot residuals against fitted
plot(full.lmod$fitted.values, sqrt(abs(full.lmod$residuals)), xlab="Fitted", ylab=expression(sqrt(hat(epsilon))))
abline(h=0)

#--------------------Normality--------------------

#Checking normality with QQ-plot
qqnorm(full.lmod$residuals, ylab="Residuals", main="")
qqline(residuals(full.lmod))

#Checking normality with histogram of residuals
hist(residuals(full.lmod), xlab="Residuals", main="")

#Checking normality with Shapiro-Wilks test
shapiro.test(full.lmod$residuals[1:4999])

#--------------------Timeseries Diagnostics--------------------

#Checking for serial correlation 
n <- length(full.lmod$residuals)
plot(tail(full.lmod$residuals, n - 1) ~ head(full.lmod$residuals, n - 1), xlab=expression(hat(epsilon)[i]), ylab=expression(hat(epsilon)[i + 1]))

summary(lm(tail(full.lmod$residuals, n - 1) ~ head(full.lmod$residuals, n - 1)))

#--------------------Checking Leverage Points--------------------
hatv <- hatvalues(full.lmod)
halfnorm(hatv, labs=data$County, ylab="Leverages")

#--------------------Checking for Outliers--------------------
stud <- rstudent(full.lmod)
data[which.max(stud),]

```
Errors appear to be approximately constant, by examination of the residuals vs. fitted. However, the histogram, QQ-plots, and Shapiro-Wilks indicate that the data is strongly non-normal. Data transformations are clearly required. Additionally, residuals display moderate serial correlation (R2 = 0.45 for a serial regression). Generalized least squares can resolve this issue. 

Furthermore, we find that Pinal County (Arizona) represents an extreme outlier on 7/29/2016 (from its large studentized error). Its removal, however, does not significantly affect the model, since the dataset is large.

```{r Checking for multicollinearity}

#Generating correlation matrix 
cor(data[,c("AQI", "HFC", "Other.GHG", "SF6", "Stationary.Combustion", "Biogenic.CO2", "HFE", "NF3", "PFC", "Short.Lived.Compounds", "Temperature", "pp_consumed_MMBtu", "CO2", "Income", "Nitrous.Oxide", "Population", "Total.Emissions", "pp_net_gen_MWh", "Methane", "Other.Fluorane")])

#Removing significantly collinear columns
data_nocol <- data[,-which(names(data) %in% c("pp_consumed_MMBtu","Total.Emissions"))]

#New fit
full.lmod.nocol <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + HFC + Other.Fluorane + 
    Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + Temperature +  
    Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_nocol)

```


```{r Data Transformations}

#Making response AQI positive by adding a small constant (1) -- can change later
data$AQI <- data$AQI + 1

#Getting model Likelihoods on a range of parameters
bx <- boxcox(full.lmod, plotit=T, lambda=seq(0.0, 0.5, by=0.001))

#Getting best boxcox parameter -- lambda ~ 0.5
lambda <- bx$x[which.max(bx$y)]

#Transforming model accordingly
full.lmod.T <- lm(formula = AQI ^ lambda ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane +  Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu 
+ Temperature + Methane + SF6 + Short.Lived.Compounds + Income + pp_net_gen_MWh, data = data)

```

Although the likelihood is maximized by this value of lambda, we find that the R2 value is still small :( 


```{r Generalized Least Squares}
full.glmod <- gls(AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + 
+ Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh,  data = data)
```
