---
title: "aqs"
author: "Ian Jiang"
date: "November 24, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# rm(list=ls())
# cat("\014") 
# setwd("C:/Users/ijiang6/Desktop")
```


```{r Loading in necessary packages}
#install.packages("MASS")
library(MASS)
#install.packages("faraway")
library(faraway)
#install.packages("nlme")
library(nlme)
library(leaps)
library(Amelia)
library(forecast)
library(pls)
```

## Loading Data

```{r loading raw data}
data_raw <- read.csv("data/data_merged_meso_weather-dropped.csv")
data_raw <- data_raw[!is.na(data_raw$AQI),]
```


```{r Loading cleaned data}
#Loading data
data <- read.csv("data/data_nomiss_specificAQI_updatedNov25.csv")
#Dropping missing values in response
data <- data[!is.na(data$AQI),]
#Assigning date from year, month, day
data$date <- as.numeric( difftime(ISOdate(data$Year, data$Month, data$Day, 0), ISOdate(2016,1,1,0), units="days"))
```

```{r remove irrelavant columns}
data_rel <- subset(data, select = c("AQI","Nitrous.Oxide", "NF3", "Other.GHG", "Total.Emissions", "HFC", "Other.Fluorane", "Biogenic.CO2", "Population", "CO2", "PFC", "HFE", "Stationary.Combustion", "pp_consumed_MMBtu", "Temperature",  "Methane", "SF6", "Short.Lived.Compounds", "Income", "pp_net_gen_MWh"))
```

```{r split data into training and test data}
smp_size <- floor(0.80 * nrow(data))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(data)), size = smp_size)

data_train <- data_rel[train_ind, ]
data_test <- data_rel[-train_ind, ]
```

```{r}
table(data$Defining.Parameter)
```

Based on the above distribution, we decided to build separate models based on "Defining Parameter" value.

```{r Load Data}
data_ozone <- data[data$Defining.Parameter == 'Ozone',]
data_PM25  <- data[data$Defining.Parameter == 'PM2.5',]
data_PM10  <- data[data$Defining.Parameter == 'PM10',]
data_SO2   <- data[data$Defining.Parameter == 'SO2',]
data_NO2   <- data[data$Defining.Parameter == 'NO2',]
data_CO    <- data[data$Defining.Parameter == 'CO',]
```

Note- update this code to split data for each dataset!!

```{r split each data into training and test data}
# smp_size <- floor(0.80 * nrow(data))
# 
# ## set the seed to make your partition reproducible
# set.seed(123)
# train_ind <- sample(seq_len(nrow(data)), size = smp_size)
# 
# data_train <- data[train_ind, ]
# data_test <- data[-train_ind, ]
```


## Visualizing data to find out outliers/data issues

```{r check missing data}
sapply(data_rel,function(x) sum(is.na(x)))
```

```{r check missing vs observed}
missmap(data, main = "Missing values vs observed")
```

```{r}
temp.lmod <- lm(formula = AQI ~ Temperature, data=data)
summary(temp.lmod)
```

```{r}
plot(data$AQI, data$Temperature)
```

```{r}
plot(data_raw$AQI, data_raw$Temperature)
```


```{r}
#data[data$County == 'iberville',]
# data_raw[data_raw$Temperature < -30,]
```


## Cleaning Data 

```{r}
# data$State.Code <- as.factor(data$State.Code)
```

## Fitting full model

```{r Full model}
lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_train)

summary(lmod)

```

```{r}
mean(data_train$AQI)
```

```{r}
pred_conf_int <- predict(lmod, new=data_test, interval="prediction")
# pred_conf_int
mean(pred_conf_int[,1])
max(pred_conf_int[,3])
min(pred_conf_int[,2])
max(pred_conf_int[,3]-pred_conf_int[,2])
```

```{r Fitting full model with County} 
#Fitting full model
full.lmod <- lm(formula = AQI ~ Nitrous.Oxide + County + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income + pp_net_gen_MWh, data = data_train)

```

```{r}
# summary(full.lmod)
summary(full.lmod)$r.squared
```

## Experimenting with states and county as factors

```{r States Model}
states.lmod <- lm(formula = AQI ~ Nitrous.Oxide + State + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_train)

summary(states.lmod)
```


```{r Region Model}
region.lmod <- lm(formula = AQI ~ State + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_train)

summary(region.lmod)
```

## Prediction accuracy

```{r check prediction accuracy}
AQI_pred <- predict(full.lmod, data_test)
accuracy(AQI_pred, data_test$AQI)
```

```{r}
AQI_pred <- predict(lmod, data_test)
accuracy(AQI_pred, data_test$AQI)
```

## Trying out stepwise regression
PS - check this, it gives error!!!
```{r Stepwise Regression}
# step.mod = regsubsets(AQI ~., data = data_train, really.big = T, method='forward')
```

## PCR

```{r PCR}
pcr.lmod <- pcr(AQI~., data=data_rel, scale=T, ncomp = 10, validation = "CV")
```

## Ridge/Lasso Regression
```{r}
# ridge.lmod <- lm.ridge(AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income + pp_net_gen_MWh, data=data_train, lambda = seq(0,1,1e-2))
# 
# select(ridge.lmod)
# summary(ridge.lmod)
```

## Different models based on Defining Parameter

### Defining Parameter = Ozone

```{r ozone full}
ozone.lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_ozone)

summary(ozone.lmod)
```

```{r ozone with county}
ozone_geo.lmod <- lm(formula = AQI ~ County + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_ozone)

summary(ozone_geo.lmod)
```

```{r ozone stepwise}
step.mod = regsubsets(AQI ~., data = data_ozone, really.big = T, method='forward')
```

Remove predictors based on R2 value-

```{r ozone with relevant factors}
ozone2.lmod <- lm(formula = AQI ~ Biogenic.CO2 + Population +  Stationary.Combustion + pp_consumed_MMBtu + Temperature  + Income +  pp_net_gen_MWh, data = data_ozone)
summary(ozone2.lmod)
```

```{r ozone with relevant factors and County}
ozone_geo2.lmod <- lm(formula = AQI ~ County+ Biogenic.CO2 + Population +  Stationary.Combustion + pp_consumed_MMBtu + Temperature  + Income +  pp_net_gen_MWh, data = data_ozone)
summary(ozone_geo2.lmod)
```

### Defining Parameter = PM.25

```{r pm25 full}
pm25.lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + 
Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature + Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_PM25)

summary(pm25.lmod)
```


```{r pm25 with County}
pm25_geo.lmod <- lm(formula = AQI ~ County + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + 
Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature + Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_PM25)

summary(pm25_geo.lmod)
```

### Defining Parameter = PM.10

```{r pm10 full}
pm10.lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_PM10)

summary(pm10.lmod)
```

```{r pm10 with county}
pm10_geo.lmod <- lm(formula = AQI ~ County + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_PM10)

summary(pm10_geo.lmod)
```

### Defining Parameter = so2

Note - commented out temporarilyfor new data - since it fails "Error in lm.fit(x, y, offset = offset, singular.ok = singular.ok, ...) : 
  0 (non-NA) cases"
  
```{r so2 full}
so2.lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_SO2)

summary(so2.lmod)
```

```{r so2 with county}
so2_geo.lmod <- lm(formula = AQI ~ County + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_SO2)

summary(so2_geo.lmod)
```

### Defining Parameter = no2

```{r no2 full}
no2.lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_NO2)

summary(no2.lmod)
```

```{r no2 with county}
no2_geo.lmod <- lm(formula = AQI ~ County + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_NO2)

summary(no2_geo.lmod)
```

### Defining Parameter = co

```{r co full}
co.lmod <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_CO)

summary(co.lmod)
```

```{r co with county}
co_geo.lmod <- lm(formula = AQI ~ County + Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_CO)

summary(co_geo.lmod)
```


```{r Diagnostics}
#--------------------Constant Variance--------------------
#Plot residuals against fitted
plot(full.lmod$fitted.values, sqrt(abs(full.lmod$residuals)), xlab="Fitted", ylab=expression(sqrt(hat(epsilon))))
abline(h=0)

#--------------------Normality--------------------
#Checking normality with QQ-plot
qqnorm(full.lmod$residuals, ylab="Residuals", main="")
qqline(residuals(full.lmod))

#Checking normality with histogram of residuals
hist(residuals(full.lmod), xlab="Residuals", main="")

#Checking normality with Shapiro-Wilks test
shapiro.test(full.lmod$residuals[1:4999])

#--------------------Timeseries Diagnostics--------------------
#Checking for serial correlation 
n <- length(full.lmod$residuals)
plot(tail(full.lmod$residuals, n - 1) ~ head(full.lmod$residuals, n - 1), xlab=expression(hat(epsilon)[i]), ylab=expression(hat(epsilon)[i + 1]))
summary(lm(tail(full.lmod$residuals, n - 1) ~ head(full.lmod$residuals, n - 1)))

#--------------------Checking Leverage Points--------------------
hatv <- hatvalues(full.lmod)
halfnorm(hatv, labs=data$County, ylab="Leverages")

#--------------------Checking for Outliers--------------------
stud <- rstudent(full.lmod)
data[which.max(stud),]

```

Errors appear to be approximately constant, by examination of the residuals vs. fitted. However, the histogram, QQ-plots, and Shapiro-Wilks indicate that the data is strongly non-normal. Data transformations are clearly required. Additionally, residuals display moderate serial correlation (R2 = 0.45 for a serial regression). Generalized least squares can resolve this issue. 

Furthermore, we find that Pinal County (Arizona) represents an extreme outlier on 7/29/2016 (from its large studentized error). Its removal, however, does not significantly affect the model, since the dataset is large.

```{r Checking for multicollinearity}
#Generating correlation matrix 
cor(data[,c("AQI", "HFC", "Other.GHG", "SF6", "Stationary.Combustion", "Biogenic.CO2", "HFE", "NF3", "PFC", "Short.Lived.Compounds", "Temperature", "pp_consumed_MMBtu", "CO2", "Income", "Nitrous.Oxide", "Population", "Total.Emissions", "pp_net_gen_MWh", "Methane", "Other.Fluorane")])

#Removing significantly collinear columns
data_nocol <- data[,-which(names(data) %in% c("pp_consumed_MMBtu","Total.Emissions"))]

#New fit
full.lmod.nocol <- lm(formula = AQI ~ Nitrous.Oxide + NF3 + Other.GHG + HFC + Other.Fluorane + 
    Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + Temperature +  
    Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh, data = data_nocol)
```


NOTE - gives an error with non-positive values for AQI. Fix it maybe!!

```{r Data Transformations}

# #Making response AQI positive by adding a small constant (1) -- can change later
# data$AQI <- data$AQI + 1
# 
# #Getting model Likelihoods on a range of parameters
# bx <- boxcox(full.lmod, plotit=T, lambda=seq(0.0, 0.5, by=0.001))
# 
# #Getting best boxcox parameter -- lambda ~ 0.5
# lambda <- bx$x[which.max(bx$y)]
# 
# #Transforming model accordingly
# full.lmod.T <- lm(formula = AQI ^ lambda ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane +  Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu 
# + Temperature + Methane + SF6 + Short.Lived.Compounds + Income + pp_net_gen_MWh, data = data)
```

Although the likelihood is maximized by this value of lambda, we find that the R2 value is still small :( 


```{r Generalized Least Squares}
full.glmod <- gls(AQI ~ Nitrous.Oxide + NF3 + Other.GHG + Total.Emissions + HFC + Other.Fluorane + 
+ Biogenic.CO2 + Population + CO2 + PFC + HFE + Stationary.Combustion + pp_consumed_MMBtu + Temperature +  Methane + SF6 + Short.Lived.Compounds + Income +  pp_net_gen_MWh,  data = data)
```

THE END!!!